% the acknowledgments section

% \newthought{Lorem ipsum dolor sit amet}, 

First, I would like to thank my advisor Patrick Baudisch. He shaped my thinking about research early on and encouraged be to big, wild and far ahead. While we might not always achieve this, it is certainly great to always keep this as the goal. His contribution of ideas for valuable research directions is one of the reasons, this thesis exists and for that I am incredibly thankful. (The other reason is hard work.)

I would also like to thank Michael Haller, who advised my master thesis and with that introduced me to the field of HCI. Without him I would certainly not be here finishing my PhD. I would also like to thank Mark Hancock and Stacey Scott, who hosted me for 5 months and who I had wonderful discussion with that further shaped my future path in research. 

Without my colleagues at the HPI lab, I could not have done my work. I want to thank Pedro Lopes, who always took the time to give me deep and insightful feedback, whenever I asked. He also always had an open ear when things got tough. I appreciate this a ton, thank you! Robert Kovacs can also not go unmentioned, as we shared an office, laughter and insightful discussions about each others projects. I want to thank Stefanie Müller for encouraging me, I always felt like she believes in me. I want to thank the entire lab: Mirela Alistar, Thijs Roumen, Lung-Pan Cheng, Oliver Schneider Sebastian Marwecki, Jotaro Shigeyama; and former lab-mates Dominik Schmidt and Tim Chen for discussions and fun times! I was also lucky to work with many wonderful students, in particular I want to mention Ludwig Wall and Johannes Frohnhofen.

Many people at HPI helped create a productive evironment and cannot go unmentioned here. I want to thank the lab's administrative assistant Frau Tholen for her support. Furthermore, I like to thank Daniela Roick-Schütze, the HPI research school that I was a part of for 4 years (Prof. Andreas Polze and Sabine Wagner), our IT-support (especially Andy Weber) and Christian from the security team.

I also want to shout out a special thanks to my friend Philipp Herholz, who was never tired of answering my technical questions. I also would like to mention Marc Alexa and Felix Naumann, who I occasionally bugged with questions about career and who patiently bared with me. 

I want to thank my parents and my sister for their support. Although my parents came as immigrants to Austria and had though jobs, they always encouraged me to study whatever I wanted to study and to strive for the best. Thank you for not setting borders!

Lastly, I have to thank my wonderful husband, love of my life, best friend, and even colleague David. Everyone in their doctoral studies knows that it can be hard sometimes. Without you, I could not have done this. You are my rock, my save haven, you are the one who believes in me when no one else (not even me) does. Thank you for your love who made me a happier, optimistic and more generous person. You definitely made me be the best version of myself. Everyday I am looking forward to our future together.